Hydras, Ordinals \& Co. \\ A library in Coq of entertaining formal mathematics

### 2021-09-27 doc/JFLA2022/paper.tex
minor corrections; add author field for the HOC book
add a TODO in the Background section

\begin{itemize}
\item formalizations of mathematics in proof assistants, such as Coq, continually grow in size and scope
\begin{itemize}
\item the Mathematical Components library and the four-color theorem and the odd order theorem in Coq
\item the Mathlib library for Lean
\end{itemize}
\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases
\item Lean and Mathematical Components have accompanying books, which are written mainly to onboard new library users
\item book authors faced with several design choices: literate documents, using tools to keep the book and the proof assistant synchronized
\end{itemize}

@@

\begin{figure}[h]
 \noindent
 For any $\alpha$ in $[0,\epsilon_0)$ and any $i$ in $\mathbb{N}$,
   \begin{itemize}
\item If $\alpha=0$, then $H'_\alpha (i)= i$.
\item If $\alpha=\beta+1$, then 
$H'_\alpha(i)=H'_\beta(i+1)$.
\item If $\alpha$ is a limit ordinal, then 
$H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(i)$.
\end{itemize}
\vspace{4pt}
\inputsnippets{HprimeDef}
\caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
\label{fig:Hprime}
\end{figure}

@@

Using $H'$ equations as rewrite rules, we can study a concrete example of hydra battle. Let us start with the hydra of figure~\ref{fig:start} and $0$ as initial step.
By a combination of rewriting and induction, we prove that the number of steps of the considered battle has at least $2^{10^{21}}$ binary digits.

  \caption{The hydra corresponding to the ordinal $\omega^3+3$}
  \label{fig:start}
\end{figure}

>>>>

\begin{itemize}
\item formalizations of mathematics in proof assistants, such as Coq, continually grow in size and scope
\begin{itemize}
\item the Mathematical Components library and the four-color theorem and the odd order theorem in Coq
\item the Mathlib library for Lean
\end{itemize}
\item to continue to be useful as examples and reusable libraries, formalizations must be well \emph{documented} and \emph{maintained} to ensure they work with new proof assistant releases \TODO[Pierre]{add evolution of styles of proof and formalization (not limited to Coq litterature)? } 
\item Lean and Mathematical Components have accompanying books, which are written mainly to onboard new library users
\item book authors faced with several design choices: literate documents, using tools to keep the book and the proof assistant synchronized
\end{itemize}

@@

Using $H'$s equations as rewrite rules, we can study a concrete example of hydra battle. Let us start with the hydra of figure~\ref{fig:start} and $0$ as initial step.
By a combination of rewriting and induction, we prove that the number of steps of the considered battle has at least $2^{10^{21}}$ binary digits.

  \caption{The hydra associated with the ordinal $\omega^3+3$}
  \label{fig:start}
\end{figure}

\begin{figure}[h]
 \noindent
 For any $\alpha$ in $[0,\epsilon_0)$ and any $i$ in $\mathbb{N}$,
   \begin{itemize}
\item If $\alpha=0$, then $H'_\alpha (i)= i$.
\item If $\alpha$ is a limit ordinal, then 
  $H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(i)$.
  \item If $\alpha=\beta+1$, then 
$H'_\alpha(i)=H'_\beta(i+1)$.
\end{itemize}
\vspace{4pt}
\inputsnippets{HprimeDef}
\caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
\label{fig:Hprime}
\end{figure}

@@
